---
title: "AK_TS_Cleaning"
author: "Matthew Ross"
date: "2/5/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}

library(signal)
library(xts)
library(dygraphs)
library(tidyverse)
library(RcppRoll)
library(lubridate)

setwd('~/git/macrosheds/qa_experimentation/')

knitr::opts_chunk$set(echo = TRUE, warning = F, message = F, 
                      comment = F)

all_data <- read_csv('data/anna_testdata.csv',
               col_types = c('Tnnncnnnin')) %>%
    dplyr::filter(datetime < mdy_hms('10-06-2016 00:00:00')) %>%
    select(-SC, -SC_filter) #what filter was used already?


```



## Butter from signal package

Notes - Really strong performance in net removal of 1 us/cm jumps but preserves too much wiggle.
Also has head/tail artifacts that can't be addressed, it seems.

```{r}


filtered <- all_data %>%
  mutate(SC_butter_noise = signal::filter(signal::butter(1,1/200,type='high'),
                                    all_data$SC_filled) %>%
           as.numeric(),
         SC_butter = SC_filled - SC_butter_noise)


all_xts <- xts(filtered %>%
                 select(starts_with('SC')),
               order.by=all_data$datetime)

dygraph(all_xts)

```


## TS decomposition and STL

Works very well, keeps some amount of noise, easier to understand.
STL produces a smoother fit, but not an obviously better one:
    peaks still cut off, 3
butter filter after decomposition (either type) removes minute fluctuations, 
    but creates the same characteristic "ramp-up" effect of butterworth alone.
    in this case, that ramp up coincides for the most part with the 48 NAs left
    by decomposition. if we just accept that we're going to have to reconstitute
    the tails of our series, we lump the major failings of decomposition and
    Butterworth filtering into one. Reconstitution by imputeTS::na_seasplit should be
    pretty effective, since periodicity and trend will be the only signals remaining.
    UPDATE: the ramp-up period is considerably longer than the sequence of NAs
    at the beginning. so this doesn't work. even if the whole rampy part
    is interpolated, it's wonkytron
    

```{r}

dts <- ts(all_data$SC_filled, deltat = 1/(24/(15/60)))
decomp <- decompose(dts)
stl_decomp <- stl(dts,
                  s.window = 7, # must be odd; min 7
                  s.degree = 0, # 0 or 1
                  # t.window = ,
                  t.degree = 1, #0 or 1
                  # l.window = ,
                  l.degree = 1,
                  robust = FALSE)

flt_butter <- signal::butter(n = 1,
                             W = 1/100,
                             type = 'low')

filtered <- all_data %>%
  mutate(SC_decomp = as.numeric(decomp$trend),
         SC_stl = as.numeric(stl_decomp$time.series[, 'trend']),
         SC_decomp = imputeTS::na_interpolation(SC_decomp),
         SC_stl = imputeTS::na_interpolation(SC_stl),
         SC_decomp_butter = as.numeric(signal::filter(filt = flt_butter,
                                                      x = SC_decomp)),
         SC_stl_butter = as.numeric(signal::filter(filt = flt_butter,
                                                   x = SC_stl)),
         index = 1:nrow(.))

#restore the NAs created by decomposition, and interpolate them
nr <- nrow(filtered)
filtered[c(1:48, (nr - 48):nr), c('SC_decomp_butter', 'SC_stl_butter')] <- NA
           
filtered <- filtered %>%
    mutate(SC_decomp_butter = imputeTS::na_interpolation(SC_decomp_butter),
           SC_stl_butter = imputeTS::na_interpolation(SC_stl_butter))

all_xts <- xts(filtered %>%
                 select(starts_with('SC')),
               order.by = all_data$datetime)

dygraph(all_xts)
```


## Loess

pretty darned righteous. no unwanted bumps. no head/tail anomalies. hard to
optimize (lots of parameters), but maybe search can be automated.

```{r}

loess_mod <- loess(SC_filled ~ index, data = filtered, span = 0.02)

filtered$SC_loess <- predict(loess_mod, newdata = filtered)

all_xts <- xts(filtered %>%
                 select(starts_with('SC')),
               order.by=all_data$datetime)

dygraph(all_xts)
```


## STL

```{r}


dts <- ts(all_data$SC_filled, deltat = 1/(24/(15/60)))

         
filtered <- all_data %>%
  mutate(SC_decomp = as.numeric(d$trend)) %>%
  mutate(index = 1:nrow(.))

all_xts <- xts(filtered %>%
                 select(starts_with('SC')),
               order.by=all_data$datetime)

dygraph(all_xts)
```

## BAM

```{r}


```


## other IIR filters from signal package 

Chebyshev I, Kaiser, and Blackman filters have ripple effect. no good
Spencer takes no parameters and returns garbage
Savitzky-Golay takes a lot of parameters, but seems promising. doesn't have the
    left-tail problem

```{r}

filtered <- all_data %>%
    mutate(SC_sgolay = as.numeric(signal::sgolayfilt(x = SC_filled,
                                                     p = 1,
                                                     n = 31, #must be odd
                                                     m = 0,
                                                     ts = 1)),
           SC_spencer = as.numeric(signal::spencer(x = SC_filled)))

dygraph(xts(x = select(filtered,
                       starts_with('SC')),
            order.by = all_data$datetime))

```

## FIR filters from signal package

all are are phase-shifted and still have left tail bogusness. there might be
parameter combinations that address this, but I haven't found them. might
be a way to optimize parameter search, but it would have to balance goodness
of fit against complexity

```{r}
#FIR filters (order must be even for highpass and bandstop type)
order <- 10
cutoff <- 0.0001

flt_hamming <- signal::fir1(n = order,
                    w = cutoff,
                    type = 'low',
                    window = hamming(order + 1))
flt_hanning <- signal::fir1(n = order,
                    w = cutoff,
                    type = 'low',
                    window = hanning(order + 1))
flt_bartlett <- signal::fir1(n = order,
                    w = cutoff,
                    type = 'low',
                    window = bartlett(order + 1))
ftl_boxcar <- signal::fir1(n = order,
                    w = cutoff,
                    type = 'low',
                    window = boxcar(order + 1))
ftl_triang <- signal::fir1(n = order,
                    w = cutoff,
                    type = 'low',
                    window = triang(order + 1))
ftl_gausswin <- signal::fir1(n = order,
                    w = cutoff,
                    type = 'low',
                    window = gausswin(order + 1))
ftl_blackman <- signal::fir1(n = order,
                    w = cutoff,
                    type = 'low',
                    window = blackman(order + 1))
ftl_flattopwin <- signal::fir1(n = order,
                    w = cutoff,
                    type = 'low',
                    window = flattopwin(order + 1))

filtered <- all_data %>%
    mutate(SC_hamming = as.numeric(signal::filter(filt = flt_hamming, x = SC_filled)),
           SC_hanning = as.numeric(signal::filter(filt = flt_hanning, x = SC_filled)),
           SC_bartlett = as.numeric(signal::filter(filt = flt_bartlett, x = SC_filled)),
           SC_boxcar = as.numeric(signal::filter(filt = ftl_boxcar, x = SC_filled)),
           SC_triang = as.numeric(signal::filter(filt = ftl_triang, x = SC_filled)),
           SC_gausswin = as.numeric(signal::filter(filt = ftl_gausswin, x = SC_filled)),
           SC_blackman = as.numeric(signal::filter(filt = ftl_blackman, x = SC_filled)),
           SC_flattopwin = as.numeric(signal::filter(filt = ftl_flattopwin, x = SC_filled)))

dygraph(xts(x = select(filtered,
                       starts_with('SC')),
            order.by = all_data$datetime))

```


## Super-beefy multi-pass approach


```{r}


```
